Classification of ordinal data in deep learning: an experimental study

### 2019-03-07 paper/ieee.tex
varios

	J. de la Torre et al.~\cite{de2018weighted} proposed the use of a continuous version of the QWK loss function for the optimization algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces overfitting and training time. Also, they checked the importance of hyper-parameter tuning. First, they defined QWK metric as follows:
	\begin{equation}
	\textcolor{blue}{\text{QWK} = 1 - \frac{\sum_{i,j} \omega_{i,j} O_{i,j}}{\sum_{i,j} \omega_{i,j} E_{i,j}},}
	\end{equation}
	\textcolor{blue}{where $\omega$ is the penalization matrix, $O$ is the confusion matrix and $E$ is the normalized outer product between the prediction and the true vector.}

	\textcolor{blue}{Then, they provided the QWK loss ($\text{QWK}_l$) definition below:}
	\begin{equation}
	\textcolor{blue}{\text{QWK}_l = 1 - \frac{\sum_{k=1}^N \sum_{c=1}^C \omega_{t_k, c} P_c(X_k)}{\sum_{i=1}^C \frac{N_i}{N} \sum_{j=1}^C ( \omega_{i,j} \sum_{k=1}^N P_j(X_k) ) },}
	\end{equation}
	\textcolor{blue}{where $X_k$ is the input data of the $k$th sample, $N$ is the number of samples, $N_i$ is the number of samples of the $i$th class and $P_c(X_k)$ is the conditional probability that the $k$th sample belongs to class $c$ given that the true class is $t_k$.}

@@

	The Proportional Odds Model (POM) arises from a statistical background and is one of the first models designed explicitly for ordinal regression~\cite{mccullagh1980regression}. It dated back to 1980 and is a member of a wider family of models recognised as Cumulative Link Models (CLM)~\cite{agresti2010analysis}. CLMs predict probabilities of groups of contiguous categories, taking the ordinal scale into account. In this way, cumulative probabilities $P(y \prec C_q |\mathbf{x})$ are estimated, which can be directly related to standard probabilities:
	\begin{equation}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = P(y = \mathcal{C}_1 | \mathbf{x}) + ... + P(y = \mathcal{C}_q | \mathbf{x}),
	\end{equation}
	\begin{equation}
		P(y = \mathcal{C}_q | \mathbf{x}) = P(y \preceq \mathcal{C}_q | \mathbf{x}) - P(y \preceq \mathcal{C}_{q-1} | \mathbf{x}),
	\end{equation}
	with $q = 2, ..., Q-1$, and considering that $P(y = \mathcal{C}_1 | \mathbf{x}) = P(y \preceq \mathcal{C}_1 | \mathbf{x})$ and $P(y \preceq \mathcal{C}_Q | \mathbf{x}) = 1$.

	\textcolor{blue}{The model is inspired by the notion of a latent variable, where $f(x)$ represents a one-dimensional mapping obtained from the output of the last layer, which has only one neuron.} The decision rule $r: \mathcal{X} \rightarrow \mathcal{Y}$ is not fitted directly, but stochastic ordering of space $\mathcal{X}$ is satisfied by the following general model form \cite{herbrich2000large}:
	\begin{equation}
		g^{-1}(P(y \preceq \mathcal{C}_q | \mathbf{x})) = b_q - \textcolor{blue}{f(\mathbf{x})}, \quad q = 1, ..., Q-1,
	\end{equation}
	where $g^{-1} : [0,1] \rightarrow (-\infty, +\infty)$ is a monotonic function often termed as the inverse link function, and $b_q$ is the threshold defined for class $\mathcal{C}_q$. Consider the latent variable $y^* = f(\mathbf{x})^* = \textcolor{blue}{f(\mathbf{x})} + \epsilon$, where $\epsilon$ is the random variable of the error. The most common choice for the probability distribution of $\epsilon$ is the logistic function (which is the default function for POM). Label $\mathcal{C}_q$ is predicted if and only if $f(\mathbf{x}) \in [b_{q-1}, b_q]$, where the function $f$ and $\mathbf{b} = (b_0, b_1, ..., b_{Q-1}, b_Q)$ are to be determined from the data. It is assumed that $b_0 = -\infty$ and $b_Q = +\infty$, so the real line defined by $f(x), x \in \mathcal{X}$, is divided into $Q$ consecutive intervals. Each interval corresponds to a category. The constraints $b_1 \le b_2 \le ... \le b_{Q-1}$ ensures that $P(y \preceq \mathcal{C}_q | \mathbf{x})$ increases with $q$ \cite{mccullagh1980regression}.
	\textcolor{blue}{This order is achieved by defining the first threshold and calculating the rest of thresholds from the first in the following form:}

@@

	These three types of links are explained below and represented in Figure \ref{fig:linkfunctions}. They all follow the same form $P(y \prec \mathcal{C}_q) = \Phi(b_q - \textcolor{blue}{f(\mathbf{x})})$ for a continuous cdf~$\Phi$.

@@

	\begin{itemize}
		\item \texttt{Logit}. \texttt{logit} link function is the function used for the Proportional Odds Model. The \texttt{logit} link is:
		\begin{equation}
		\begin{aligned}
		\text{logit}[P(y \preceq \mathcal{C}_q)] = \log\frac{P(y \preceq \mathcal{C}_q)}{1 - P(y \preceq \mathcal{C}_q)}=& \\ = b_q - \textcolor{blue}{f(\mathbf{x})}, \quad q = 1, ..., Q-1,
		\end{aligned}
		\label{eq:logit}
		\end{equation}		
		or the equivalent expression:		
		\begin{equation}
		P(y \preceq \mathcal{C}_q) = \frac{1}{1 + e^{-(b - \textcolor{blue}{f(\mathbf{x})})}}.
		\label{eq:logit2}
		\end{equation}

		\item \texttt{Probit}. \texttt{probit} link function is the inverse of the standard normal cumulative distribution function (cdf) $\Phi$. Its expression is:
		\begin{equation}
		\begin{aligned}
		&\Phi^{-1}[P(y \preceq \mathcal{C}_q)] = b_q - \textcolor{blue}{f(\mathbf{x})}, &q = 1, ..., Q-1\\
		&P(y \preceq \mathcal{C}_q) = \Phi(b_q - \textcolor{blue}{f(\mathbf{x})}), &q = 1, ..., Q-1,
		\end{aligned}
		\label{eq:probit}
		\end{equation}		
		which can also be expressed as:
		\begin{equation}
		P(y \preceq \mathcal{C}_q) = \int_{-\infty}^{b_q - \textcolor{blue}{f(\mathbf{x})}} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2}x^2} \mathrm{d}x.
		\label{eq:probit2}
		\end{equation}

		\item \textit{Complementary log-log}. Like the \texttt{logit} and the \texttt{probit} transformation, the complementary log-log transformation takes a response that is restricted to the $(0,1)$ interval and converts it into something in the $(-\infty, +\infty)$ interval. Complementary log-log expression is:
		\begin{equation}
		\begin{aligned}
		\log[-\log[1 - P(y \preceq \mathcal{C}_q)]] =&\\
		= b_q - \textcolor{blue}{f(\mathbf{x})}, \quad q = 1, ..., Q-1,&
		\end{aligned}
		\label{eq:cloglog}
		\end{equation}
		that is:
		\begin{equation}
		P(y \preceq \mathcal{C}_q) = 1 - e^{-e^{b_q - \textcolor{blue}{f(\mathbf{x})}}}, \quad q = 1, ..., Q-1.
		\label{eq:cloglog2}
		\end{equation}
	\end{itemize}

	\texttt{Logit} and \texttt{probit} links are symmetric:	
	\begin{equation}
	\text{link}[P(y \preceq \mathcal{C}_q)] = -\text{link}[1 - P(y \preceq \mathcal{C}_q)],
	\end{equation}
	which means that the response curve for $P(y \preceq \mathcal{C}_q)$ is symmetric around the point $P(y \preceq \mathcal{C}_q) = 0.5$, i.e. $P(y \preceq \mathcal{C}_q)$ has the same rate when approaching 0 than when approaching 1. This symmetric property can be demonstrated as follows:	
	\begin{enumerate}
		\item Let $P(y \preceq \mathcal{C}_q) \equiv p$. For the \texttt{logit} function, we have:
		\begin{equation}
		\begin{aligned}
		\text{link}(p) = \text{logit}(p) &= \log\left(\frac{p}{1-p}\right) =\\
		&= \log(p) - \log(1-p),
		\end{aligned}
		\end{equation}			
		while:			
		\begin{equation}
		\begin{aligned}
		-\text{link}(1 - p) &= -\text{logit}(1 - p) =\\
		=- \log\left(\frac{1- p}{p}\right) &= - \log(1 - p) + \log(p).
		\end{aligned}
		\end{equation}

		\item For the \texttt{probit}:		
		\begin{equation}
		\begin{aligned}
		p \equiv P(y \preceq \mathcal{C}_q) &= \Phi(b_q - \textcolor{blue}{f(\mathbf{x})}) =\\
		&= \int_{-\infty}^{b_q - \textcolor{blue}{f(\mathbf{x})}} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2}x^2} \mathrm{d}x,
		\end{aligned}
		\end{equation}
		which leads to:
		\begin{align}
		\text{probit}(p) & = \Phi^{-1}(p) = b_q - \textcolor{blue}{f(\mathbf{x})},\\
		-\text{probit}(1-p) &= \Phi^{-1}(1-p) = -b_q + \textcolor{blue}{f(\mathbf{x})},
		\end{align}
		where:		
		\begin{equation}
		1 - p = \int_{-\infty}^{-b_q + \textcolor{blue}{f(\mathbf{x})}} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2}x^2} \mathrm{d}x,
		\end{equation}
		\begin{equation}
		p = 1 - \int_{-\infty}^{-b_q + \textcolor{blue}{f(\mathbf{x})}} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2}x^2} \mathrm{d}x.
		\end{equation}
	\end{enumerate}

@@

	\begin{itemize}
		\item \textit{Diabetic Retinopathy (DR)}\footnote{https://www.kaggle.com/c/diabetic-retinopathy-detection/data}. DR is a dataset consisting of extremely high-resolution fundus image data. The training set consists of $17563$ pairs of images (where a  pair consists of a left and right eye image corresponding to a patient). In this dataset, we try to predict the correct category from five levels of diabetic retinopathy: no DR ($25810$ images), mild DR ($2443$ images), moderate DR ($5292$ images), severe DR ($873$ images), or proliferative DR ($708$ images). The images are taken in variable conditions: by different cameras,  varying conditions of illumination and different resolutions. These images come from the EyePACS dataset that was used in a Diabetic Retinopathy Detection competition hosted on the Kaggle platform. Also, this dataset was used in later works~\cite{de2018weighted,nebot2016diabetic}, applying an ordinal QWK cost function in \cite{de2018weighted} to achieve better performance. A validation set is set aside, consisting of $10\%$ of the patients in the training set. The images are resized to 128 by 128 pixels and rescaled to $[0,1]$ range. Data augmentation techniques, described in Section \ref{sect:settings}, are applied to achieve a higher number of samples. A few images of this dataset are shown in Figure \ref{fig:DRexamples}.

@@

	The best mean QWK value was obtained with the complementary log-log link function using a batch size of 10 and a learning rate of $10^{-3}$. However, the best CCR value was obtained with a batch size of 15, the \texttt{logit} link and a learning rate of $10^{-4}$. The optimal configuration depends on the metric we are analysing. In this case, as we are working with an ordinal problem, the most reliable metric is the QWK. \textcolor{red}{No se entonces si tiene sentido poner el resto de métricas???} \textcolor{blue}{Están puestas para comparar con artículos que no dan el QWK.}

@@

	\subsection{Statistical analysis}
	\label{sect:statisticalanalysis}
	In this subsection, a statistical analysis will be performed in order to obtain conclusions from the result,

@@

	We denote by $QWK_{i,j,k,l}(i=1, ..., 3; j = 1, ..., 3; k = 1, ..., 3)$ the value observed when the first factor is at the $i$-th level, the second at the $j$-th level and the third at the $k$-th level. We assume that the three factors do not act independently and therefore there exists an interaction between each pair of them and between the three factors. In this case, the observations fit:
	\begin{equation}
	\begin{aligned}
	\footnotesize
	QWK_{i,j,k,l} = & \mu + L_i + P_j + B_k + LP_{i,j} + LB_{i,k} \\& + PB_{j,k} + LPB_{i,j,k} + \epsilon_{i,j,k,l},
	\label{eq:anova3}
	\end{aligned}
	\end{equation}
	where $\mu$ is the fixed effect that is common to all the populations; $L_i$ is the effect associated with the \textit{i-th} level of the link factor (\texttt{logit}, \texttt{probit}, complementary log-log); $P_j$ is the effect associated with the \textit{j-th} level of the learning rate factor and $B_k$ is the effect associated with the \textit{k-th} level of the batch size factor. The term $LP_{i,j}$ denotes the joint effect of the presence of level $i$ of the first factor and level $j$ of the second one; this, therefore, is denominated the interaction term between $L$ and $P$ factors. The same interaction effect is appreciated on $LB_{i,k}$, $PB_{j,k}$ and $LPB_{i,j,k}$. The term $\epsilon_{i,j,k,l}$ is the influence on the result of everything that could not be assigned or of random factors. $QWK_{i,j,k,l}$ is the quadratic weighted kappa measure, the response variable used to perform the statistical analysis.

@@

	Finally, a comparison of the best \textcolor{blue}{test results} for each dataset for ordinal and nominal cases and previous works is shown in tables \ref{table:ComparisonDR} and \ref{table:ComparisonAdience}. The proposed ordinal model outperforms all the other alternatives in terms of QWK. The performance gain of CLM over the softmax reaches $16.8\%$ for Diabetic Retinopathy and $11.9\%$ for Adience dataset. The improvement of the ordinal method for Retinopathy dataset is higher than for Adience dataset. It seems that the method proposed in this work offers a more significant improvement as the given problem complexity increases.

>>>>

	J. de la Torre et al.~\cite{de2018weighted} proposed the use of a continuous version of the QWK loss function for the optimization algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces overfitting and training time. Also, they checked the importance of hyper-parameter tuning. First, they defined QWK metric as follows:
	\begin{equation}
	\textcolor{blue}{\text{QWK} = 1 - \frac{\sum\limits_{i,j} \omega_{i,j} O_{i,j}}{\sum\limits_{i,j} \omega_{i,j} E_{i,j}},}
	\end{equation}
	\textcolor{blue}{where $\omega$ is the penalization matrix (in this case, quadratic weights are considered), $O$ is the confusion matrix and $E$ is the normalized outer product between the prediction and the true vector.}

	\textcolor{blue}{Then, they provided the QWK loss ($\text{QWK}_l$) definition below:}
	\begin{equation}
	\textcolor{blue}{\text{QWK}_l = \frac{\sum\limits_{k=1}^N \sum\limits_{q=1}^Q \omega_{t_k, q} P(y = \mathcal{C}_q | \mathbf{x}_k)}{\sum\limits_{i=1}^Q \frac{N_i}{N} \sum\limits_{j=1}^Q ( \omega_{i,j} \sum\limits_{k=1}^N P(y = \mathcal{C}_j | \mathbf{x}_k))},}
	\end{equation}
	\textcolor{blue}{where $\text{QWK}_l \in [0,2]$, $\mathbf{x}_k$ and $t_k$ are the input data and the real class of the $k$-th sample, $Q$ is the number of classes, $N$ is the number of samples, $N_i$ is the number of samples of the $i$th class, $P(y = \mathcal{C}_q | \mathbf{x}_k)$ is the probability that the $k$th sample belongs to class $\mathcal{C}_q$ and $\omega_{i,j}$ are the elements of the penalization matrix. Generally, $\omega_{i,j} = \frac{|i-j|^n}{(C-1)^n}$, where $\omega_{i,j} \in [0,1]$. In this case, the $\text{QWK}_l$ is a function to be minimized while the QWK metric must be maximized.}

@@

	The Proportional Odds Model (POM) arises from a statistical background and is one of the first models designed explicitly for ordinal regression~\cite{mccullagh1980regression}. It dated back to 1980 and is a member of a wider family of models recognised as Cumulative Link Models (CLM)~\cite{agresti2010analysis}. CLMs predict probabilities of groups of contiguous categories, taking the ordinal scale into account. In this way, cumulative probabilities $P(y \prec C_q |\mathbf{x})$ are estimated, which can be directly related to standard probabilities:
	\begin{align}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) &= P(y = \mathcal{C}_1 | \mathbf{x}) + ... + P(y = \mathcal{C}_q | \mathbf{x}),\\
		P(y = \mathcal{C}_q | \mathbf{x}) &= P(y \preceq \mathcal{C}_q | \mathbf{x}) - P(y \preceq \mathcal{C}_{q-1} | \mathbf{x}),
	\end{align}
	with $q = 2, ..., Q-1$, and considering that $P(y = \mathcal{C}_1 | \mathbf{x}) = P(y \preceq \mathcal{C}_1 | \mathbf{x})$ and $P(y \preceq \mathcal{C}_Q | \mathbf{x}) = 1$.

	\textcolor{blue}{The model is inspired by the notion of a latent variable, where $f(\textbf{x})$ represents a one-dimensional mapping obtained from the output of the last layer, which has only one neuron.} The decision rule $r: \mathcal{X} \rightarrow \mathcal{Y}$ is not fitted directly, but stochastic ordering of space $\mathcal{X}$ is satisfied by the following general model form \cite{herbrich2000large}:
	\begin{equation}
		g^{-1}(P(y \preceq \mathcal{C}_q | \mathbf{x})) = b_q - \textcolor{blue}{f(\mathbf{x})}, \quad q = 1, ..., Q-1,
	\end{equation}
	where $g^{-1} : [0,1] \rightarrow (-\infty, +\infty)$ is a monotonic function often termed as the inverse link function, and $b_q$ is the threshold defined for class $\mathcal{C}_q$. Consider the latent variable $y^* = f(\mathbf{x})^* = \textcolor{blue}{f(\mathbf{x})} + \epsilon$, where $\epsilon$ is the random variable of the error. The most common choice for the probability distribution of $\epsilon$ is the logistic function (which is the default function for POM). Label $\mathcal{C}_q$ is predicted if and only if $f(\mathbf{x}) \in [b_{q-1}, b_q]$, where the function $f$ and $\mathbf{b} = (b_0, b_1, ..., b_{Q-1}, b_Q)$ are to be determined from the data. It is assumed that $b_0 = -\infty$ and $b_Q = +\infty$, so the real line defined by $f(\textbf{x}), \textbf{x} \in \mathcal{X}$, is divided into $Q$ consecutive intervals. Each interval corresponds to a category. The constraints $b_1 \le b_2 \le ... \le b_{Q-1}$ ensures that $P(y \preceq \mathcal{C}_q | \mathbf{x})$ increases with $q$ \cite{mccullagh1980regression}.
	\textcolor{blue}{This order is achieved by defining the first threshold and calculating the rest of thresholds from the first in the following form:}

@@

	These three types of links are explained below and represented in Figure \ref{fig:linkfunctions}. They all follow the same form $P(y \preceq \mathcal{C}_q | \mathbf{x}) = \Phi(b_q - \textcolor{blue}{f(\mathbf{x})})$ for a continuous cdf~$\Phi$.

@@

	\begin{itemize}
		\item \texttt{Logit}. \texttt{logit} link function is the function used for the Proportional Odds Model. The \texttt{logit} link is:
		\begin{equation}
		\begin{aligned}
		\text{logit}[P(y \preceq \mathcal{C}_q | \mathbf{x})] = \log\frac{P(y \preceq \mathcal{C}_q | \mathbf{x})}{1 - P(y \preceq \mathcal{C}_q | \mathbf{x})}=& \\ = b_q - \textcolor{blue}{f(\mathbf{x})}, \quad q = 1, ..., Q-1,
		\end{aligned}
		\label{eq:logit}
		\end{equation}		
		or the equivalent expression:		
		\begin{equation}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = \frac{1}{1 + e^{-(b - \textcolor{blue}{f(\mathbf{x})})}}.
		\label{eq:logit2}
		\end{equation}

		\item \texttt{Probit}. \texttt{probit} link function is the inverse of the standard normal cumulative distribution function (cdf) $\Phi$. Its expression is:
		\begin{equation}
		\begin{aligned}
		\Phi^{-1}[P(y \preceq \mathcal{C}_q | \mathbf{x})] = b_q - \textcolor{blue}{f(\mathbf{x})}, \quad &q = 1, ..., Q-1,\\
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = \Phi(b_q - \textcolor{blue}{f(\mathbf{x})}), \quad &q = 1, ..., Q-1,
		\end{aligned}
		\label{eq:probit}
		\end{equation}		
		which can also be expressed as:
		\begin{equation}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = \int_{-\infty}^{b_q - \textcolor{blue}{f(\mathbf{x})}} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2}x^2} \mathrm{d}x.
		\label{eq:probit2}
		\end{equation}

		\item \textit{Complementary log-log}. Like the \texttt{logit} and the \texttt{probit} transformation, the complementary log-log transformation takes a response that is restricted to the $(0,1)$ interval and converts it into something in the $(-\infty, +\infty)$ interval. Complementary log-log expression is:
		\begin{equation}
		\begin{aligned}
		\log[-\log[1 - P(y \preceq \mathcal{C}_q | \mathbf{x})]] =&\\
		= b_q - \textcolor{blue}{f(\mathbf{x})}, \quad q = 1, ..., Q-1,&
		\end{aligned}
		\label{eq:cloglog}
		\end{equation}
		that is:
		\begin{equation}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = 1 - e^{-e^{b_q - \textcolor{blue}{f(\mathbf{x})}}}, \quad q = 1, ..., Q-1.
		\label{eq:cloglog2}
		\end{equation}
	\end{itemize}

	\texttt{Logit} and \texttt{probit} links are symmetric:	
	\begin{equation}
	\text{link}[P(y \preceq \mathcal{C}_q | \mathbf{x})] = -\text{link}[1 - P(y \preceq \mathcal{C}_q | \mathbf{x})],
	\end{equation}
	which means that the response curve for $P(y \preceq \mathcal{C}_q | \mathbf{x})$ is symmetric around the point $P(y \preceq \mathcal{C}_q | \mathbf{x}) = 0.5$, i.e. $P(y \preceq \mathcal{C}_q | \mathbf{x})$ has the same rate when approaching 0 than when approaching 1. This symmetric property can be demonstrated as follows:	
	\begin{enumerate}
		\item Let $P(y \preceq \mathcal{C}_q | \mathbf{x}) \equiv p$. For the \texttt{logit} function, we have:
		\begin{equation}
		\begin{aligned}
		\text{link}(p) = \text{logit}(p) &= \log\left(\frac{p}{1-p}\right) =\\
		&= \log(p) - \log(1-p),
		\end{aligned}
		\end{equation}			
		while:			
		\begin{equation}
		\begin{aligned}
		-\text{link}(1 - p) &= -\text{logit}(1 - p) =\\
		=- \log\left(\frac{1- p}{p}\right) &= - \log(1 - p) + \log(p).
		\end{aligned}
		\end{equation}

		\item For the \texttt{probit}:		
		\begin{equation}
		\begin{aligned}
		p \equiv P(y \preceq \mathcal{C}_q | \mathbf{x}) &= \Phi(b_q - \textcolor{blue}{f(\mathbf{x})}) =\\
		&= \int_{-\infty}^{b_q - \textcolor{blue}{f(\mathbf{x})}} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2}x^2} \mathrm{d}x,
		\end{aligned}
		\end{equation}
		which leads to:
		\begin{align}
		\text{probit}(p) & = \Phi^{-1}(p) = b_q - \textcolor{blue}{f(\mathbf{x})},\\
		-\text{probit}(1-p) &= \Phi^{-1}(1-p) = -b_q + \textcolor{blue}{f(\mathbf{x})},
		\end{align}
		where:		
		\begin{equation}
		1 - p = \int_{-\infty}^{-b_q + \textcolor{blue}{f(\mathbf{x})}} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2}x^2} \mathrm{d}x,
		\end{equation}
		\begin{equation}
		p = 1 - \int_{-\infty}^{-b_q + \textcolor{blue}{f(\mathbf{x})}} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2}x^2} \mathrm{d}x.
		\end{equation}
	\end{enumerate}

@@

	\begin{itemize}
		\item \textit{Diabetic Retinopathy (DR)}\footnote{https://www.kaggle.com/c/diabetic-retinopathy-detection/data}. DR is a dataset consisting of extremely high-resolution fundus image data. The training set consists of $17563$ pairs of images (where a  pair consists of a left and right eye image corresponding to a patient). In this dataset, we try to predict the correct category from five levels of diabetic retinopathy: no DR ($25810$ images), mild DR ($2443$ images), moderate DR ($5292$ images), severe DR ($873$ images), or proliferative DR ($708$ images). The test set contains $26788$ pairs of images. These images are taken in variable conditions: by different cameras,  varying conditions of illumination and different resolutions. These images come from the EyePACS dataset that was used in a Diabetic Retinopathy Detection competition hosted on the Kaggle platform. Also, this dataset was used in later works~\cite{de2018weighted,nebot2016diabetic}, applying an ordinal QWK cost function in \cite{de2018weighted} to achieve better performance. A validation set is set aside, consisting of $10\%$ of the patients in the training set. The images are resized to 128 by 128 pixels and rescaled to $[0,1]$ range. Data augmentation techniques, described in Section \ref{sect:settings}, are applied to achieve a higher number of samples. A few images of this dataset are shown in Figure \ref{fig:DRexamples}.

@@

	The best mean QWK value was obtained with the complementary log-log link function using a batch size of 10 and a learning rate of $10^{-3}$. However, the best CCR value was obtained with a batch size of 15, the \texttt{logit} link and a learning rate of $10^{-4}$. The optimal configuration depends on the metric we are analysing. \textcolor{blue}{In this case, as we are working with an ordinal problem, the most reliable metric is the QWK. However, the rest of the metrics are also included to allow further comparisons with future works.}

@@

	\subsection{Statistical analysis}
	\label{sect:statisticalanalysis}
	In this subsection, a statistical analysis will be performed in order to obtain conclusions from the result.

@@

	We denote by $\text{QWK}_{i,j,k,l}(i=1, ..., 3; j = 1, ..., 3; k = 1, ..., 3)$ the value observed when the first factor is at the $i$-th level, the second at the $j$-th level and the third at the $k$-th level. We assume that the three factors do not act independently and therefore there exists an interaction between each pair of them and between the three factors. In this case, the observations fit:
	\begin{equation}
	\begin{aligned}
	\footnotesize
	\text{QWK}_{i,j,k,l} = \mu & + L_i + P_j + B_k + LP_{i,j} + LB_{i,k} \\& + PB_{j,k} + LPB_{i,j,k} + \epsilon_{i,j,k,l},
	\label{eq:anova3}
	\end{aligned}
	\end{equation}
	where $\mu$ is the fixed effect that is common to all the populations; $L_i$ is the effect associated with the \textit{i-th} level of the link factor (\texttt{logit}, \texttt{probit}, complementary log-log); $P_j$ is the effect associated with the \textit{j-th} level of the learning rate factor and $B_k$ is the effect associated with the \textit{k-th} level of the batch size factor. The term $LP_{i,j}$ denotes the joint effect of the presence of level $i$ of the first factor and level $j$ of the second one; this, therefore, is denominated the interaction term between $L$ and $P$ factors. The same interaction effect is appreciated on $LB_{i,k}$, $PB_{j,k}$ and $LPB_{i,j,k}$. The term $\epsilon_{i,j,k,l}$ is the influence on the result of everything that could not be assigned or of random factors. $QWK_{i,j,k,l}$ is the quadratic weighted kappa measure, the response variable used to perform the statistical analysis.

@@

	Finally, a comparison of the best results for each dataset for ordinal and nominal cases and previous works is shown in tables \ref{table:ComparisonDR} and \ref{table:ComparisonAdience}. \textcolor{blue}{All the results are given for the test set, except those from \cite{de2018weighted} (DR dataset), because the authors only provided validation results for $128\times 128$ images (however, validation results are usually better than test results).} The proposed ordinal model outperforms all the other alternatives in terms of QWK. The performance gain of CLM over the softmax reaches $16.8\%$ for Diabetic Retinopathy and $11.9\%$ for Adience dataset. The improvement of the ordinal method for Retinopathy dataset is higher than for Adience dataset. It seems that the method proposed in this work offers a more significant improvement as the given problem complexity increases.
